JD notes on effects

### 2022-03-23 manuscript/vareffects_plosone.tex
Autocommit

Several quantities of interest may be derived from regression models. The first one is the coefficient estimates. Others are \emph{input variables},  \emph{predictors}, \emph{model matrix}, \emph{predicted values} and \emph{marginal effects}. Input variables are the variables that were measured (may be transformed), while predictors are the terms that are entered in the model. Hence, predictors encompass the main effects, but also polynomials of input variables and interaction terms (for example in polynomials and splines) \cite{schielzeth2010simple}. Model matrix refers to the design matrix whose rows include all combination of variables appearing in the interaction terms, along with the typical values of the focal and non-focal predictors.

>>>>

Several quantities of interest may be derived from regression models. The first one is the coefficient estimates. Others are \emph{input and model variables}, \emph{model matrix}, \emph{model center}, \emph{reference point}, \emph{anchor}, \emph{predictors values} and \emph{marginal effects}. In particular: input variables are scientific variables underlying an inference or exploration. The focal predictor we use for effects or predictions is an input variable \cite{schielzeth2010simple}; model matrix refers to the design matrix whose rows include all combination of variables appearing in the interaction terms, along with the typical values of the focal and non-focal predictors; model variables are variables that represent columns in the model matrix. Each input variable will correspond to one or more model variables. In particular, variables with more than two categories, or variables modeled with a spline or polynomial response, will correspond to more than one input variable; model center is a point corresponding to a column-wise mean of the model matrix (the mean of one or more model variables). The center point for a set of model variables corresponding to an input variable may not represent a possible value of the input variable; reference point is a (or are) value (or values) chosen for non-focal predictors, when estimating effects. Typically the center point, but can instead be a population of quantiles or observations; and lastly, an anchor is the value chosen for the focal predictor when estimating effect confidence intervals (anchor choice does not affect the estimate). Typically chosen as the center point of the model variables corresponding to the focal input variable.
